The reference model approach in feature selection problems \vspace{.1in}

### 2019-07-16 paper/ref_approach.tex
start editing Section 3

In this Section we show the benefits of the use of a reference model using as a benchmark the normal means estimation problem, which consists in estimating the vector of means, usually sparse, of a normal vector of observations. Let us call $p$ the dimensionality of the vector of means and let $\{z_{j}\}_{j=1}^{p}$ be the vector of observations of the random variables $\{Z_{j}\}_{j=1}^{p}$, the normal means problem consists in estimating the latent variables $\{\theta_{j}\}_{j=1}^{p}$ of the following model:
\
\begin{equation}\label{eq:normal_means_problem}
Z_{j}|\theta_{j},\sigma\overset{ind}{\sim}N(\theta_{j},\sigma^{2}) \quad j=1,..,p
\end{equation}
Note that it is equivalent to a linear regression where the design matrix is the identity one. This formulation can be retrieved in different ways, a common example is given by microarray data, where a large set of genes are usually tested in two groups of patient labelled as positive or negative to some disease. The objective of the analysis is to spot the whole subset of statistically relevant genes to the disease and one possible way to proceed is to compute the two-samples $t$-statistics for each gene and after combining it with the cumulative density function of the standard distribution, the resulting data can be used in the problem formulation \eqref{eq:normal_means_problem}. For further details see the examples in \cite{paper:efron, efron2012large}. 
\\
In our examples we retrieve the normal means problem from the sample correlation using the Fisher $z$-transformation approximation \citep{hawkins1989using}. Suppose to have a continuous target random variable $Y$ and a set of $p$ continuous covariates $\{X_{j}\}_{j=1}^{p}$ and let us call $\rho_{j}=Cor(Y,X_{j})$. Suppose to observe $n$ statistical units and define $r_{j}$ the sample correlation between the observations of the target variables $\{y_{i}\}_{i=1}^{n}$ and the j-th covariate $\{x_{ij}\}_{i=1}^{n}$. Let finally refer to the function $\text{tanh}^{-1}(\cdot)$ as $T_{F}(\cdot)$. It holds the following modelling approximation (assuming each pair $(Y,X_{j})$ normally distributed):
\
\begin{equation} \label{eq:fisher_transformation}
T_{F}(r_{j})\overset{ind}{\sim} N(T_{F}(\rho_{j}),\frac{1}{n-3}) \quad j=1,..,p
\end{equation}
Therefore rescaling the quantities $T_{F}(r_{j})$ by $\sqrt{n-3}$ and referring to the results as variables $z_{j}$, we find again formulation \eqref{eq:normal_means_problem} with unit variance:
\
\begin{equation} \label{eq:normal_means_problem2}
Z_{j}|\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \quad j=1,..,p
\end{equation}
Note that now the quantity of interest $\theta_{j}$ stands for $\sqrt{n-3}\,T_{F}(\rho_{j})$. There are different ways to proceed with the selection from the normal means problem, we consider in our comparison both frequentist and Bayesian state-of-the-art methods. From a full-Bayesian perspective, a Bayesian regression model is fitted to the normal means problem typically using a sparsifying prior as in \cite{paper:dirichlet_laplace}, \cite{bhadra2017horseshoe+} and \cite{johnstone2004needles}. In case of continuous priors, as the Dirichlet-Laplace (DL) and the horseshoe+ priors, there is not a canonical way to complete the selection since the posterior probability of an exactly zero signal is always zero; in their experiments \cite{paper:dirichlet_laplace} use k-means clustering over the posterior samples using the DL prior, whereas \cite{bhadra2017horseshoe+} focus the analysis on the obtained shrinkage and the correct estimation of the sparse signals using the horseshoe+ prior. We include in our experiments the DL prior and the regularised horseshoe (RHS) prior \citep{paper:rhs}, which has shown in our opinion better scalability with respect to the horseshoe+, comparing the shrinkage and the correct estimation of the sparse signals using or not the reference model approach. The quantity of interests that we consider are the average SSE (sum of square errors) and the average SESE (sum of expected square errors), defined as follows:
\
\begin{align}
\text{SSE}&=\frac{1}{n-3}\sum_{j=1}^{p}(\hat{\theta}_{j} - \theta^{0}_{j})^{2} \label{eq:SSE} \\
\text{SESE}&=\frac{1}{n-3}\sum_{j=1}^{p}\int_{\mathbb{R}}(\theta_{j}-\theta^{0}_{j})^{2}p(\theta_{j}|z_{1},..,z_{p})d\theta_{j} \\
&=\frac{1}{n-3}\sum_{j=1}^{p}\mathbb{E}_{\theta_{j}}[(\theta_{j}-\theta^{0}_{j})^{2}|z_{1},..,z_{p}] \label{eq:SESE}
\end{align}
where $\theta_{j}^{0}=\sqrt{n-3}\,T_{F}(\rho_{j})$ is the true value, $\hat{\theta}_{j}$ is the posterior median of each parameter $\theta_{j}$ and \eqref{eq:SESE} is a sum of posterior expectations. We divide the errors by $(n-3)$ to have quantities independent from the number of statistical units. Note that the quantity SESE takes in account also the uncertainty, thus the shrinkage, of the posterior distributions, whereas the SSE measures only the bias in the point estimates. Note also that any kind of point estimate could be used to compute the SSE, here we follow \cite{paper:dirichlet_laplace} using the posterior median. The results that we show later in Section \ref{shrinkage-signal} are not sensible to such a choice. We also take into account the posterior mean for the shrinkage factor ($\kappa_{j}$) for each parameter $\theta_{j}$, which for model \eqref{eq:normal_means_problem2} with a hierarchical shrinkage prior of the type:
\begin{align}
Z_{j}|&\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \qquad j=1,..,p \\
&\theta_{j}|\tau,\nu_{j}\overset{ind}{\sim}N(0,\nu_{j}^{2}\tau^{2})
\end{align}
is defined as:
\begin{equation}
\kappa_{j}=\frac{1}{1+\tau^{2}\nu_{j}^{2}}
\end{equation}
\
and, thus, in our case it becomes:
\begin{equation}
\kappa_{j}^{\text{RHS}}=\frac{1}{1+\tau^{2}\lambda_{j}^{2}}, \qquad \kappa_{j}^{\text{DL}}=\frac{1}{1+\psi\phi_{j}^{2}\tau^{2}}
\end{equation}
respectively for the regularised horseshoe and the Dirichlet-Laplace priors. In the left hand-side expression, $\tau$ and $\lambda_{j}$ stand for the global and local shrinkage parameters of the regularised horseshoe prior \cite[notation and further details in][]{paper:rhs}, whereas on the right hand-side $\psi\phi_{j}^{2}\tau^{2}$ is the variance term of the Dirichlet-Laplace prior of parameter $\theta_{j}$, following the same notation as \cite{paper:dirichlet_laplace}.

\cite{johnstone2004needles} use a prior of spike-and-slab type \citep{paper:spike_slab_mitchell} given by a mixture of a delta distribution in zero and a heavy-tailed distribution; when the latter is a Laplace distribution they show an interesting thresholding property of the median estimator and, thus, present a complete selection procedure using their prior. We include in our comparison such method (EB.med) together with the control of the local false discovery rate \cite[loc.fdr; see][]{efron2012large, paper:efron} and a simple selection procedure using inclusion probability at 0.9 for posterior credible intervals using the regularised horseshoe prior (ci.90). All these methods provide a complete selection procedures, we compare the result using or not a reference model on top of the procedure evaluating the average false discovery rate (i.e. ratio of the number of non-relevant selected features over the number of selected features) and the average sensitivity (i.e. ratio of the number of relevant selected features over the total number of relevant features). We also provide a comparison of the stability of the selection using the measure proposed by \cite{paper:stability}. Table \ref{tab:comparison} summarises the methods used in the comparison.

\begin{table}[tp]
\scriptsize
\centering
\begin{tabular}{l|l|l|l}
  \hline
 Method & Alias & Type & Quantities of interest \\ 
  \hline
 Regularised Horseshoe & RHS & Shrinkage prior & SSE, SESE, $\kappa$ \\
 Dirichlet-Laplace & DL & Shrinkage prior & SSE, SESE, $\kappa$ \\
 Local false discovery rate & loc.fdr & Complete selection & fdr, sensitivity, stability \\
 Median estimator & EB.med & Complete selection & fdr, sensitivity, stability \\
 Credible intervals & ci.90 & Complete selection & fdr, sensitivity, stability \\
 \hline
\end{tabular}
\caption{Summary of the methods used in the comparison.}
\label{tab:comparison}
\end{table}

>>>>

In this section, we demonstrate the benefits of reference model approaches using the normal means problem as a benchmark. The normal means problem consists of estimating the (usually sparse) vector of means of a vector of normally distributed observations. Let us denote the dimensionality of the vector of means by $p$ and let $\{z_{j}\}_{j=1}^{p}$ be the vector of observations of the random variables $\{Z_{j}\}_{j=1}^{p}$. In the normal means problem, the task is to estimate the latent variables $\{\theta_{j}\}_{j=1}^{p}$ of the following model:
\
\begin{equation}\label{eq:normal_means_problem}
Z_{j}|\theta_{j},\sigma\overset{ind}{\sim}N(\theta_{j},\sigma^{2}) \quad j=1,..,p
\end{equation}
Note that it is equivalent to a linear regression where the design matrix is the identity matrix with the number of features being equal to the number of observations. This formulation can be found in practice at various places, 
for example, in the analysis of microarray data, where a large set of genes are tested in two groups of patients labeled as positive or negative to some disease (cite). The objective of the analysis is to select the subset of genes statistically relevant to the disease. One common way to proceed is to compute the two-samples $t$-statistic for every gene separately.
After normalizing these statistics, they then become the data, $Z$, in the normal means problem \eqref{eq:normal_means_problem}. For further details see the examples in \cite{paper:efron, efron2012large}.

In our examples, we retrieve the normal means problem from the sample correlations between the target and the covariates using the Fisher $z$-transformation \citep{hawkins1989using}. Suppose we have a continuous target random variable $Y$, a set of $p$ continuous covariates $\{X_{j}\}_{j=1}^{p}$, and let us denote $\rho_{j}=Cor(Y,X_{j})$. Further, suppose we have observed $n$ statistical units and define $r_{j}$ the sample correlation between the observations of the target variables $\{y_{i}\}_{i=1}^{n}$ and the $j$-th covariate $\{x_{ij}\}_{i=1}^{n}$. Finally, we refer to the Fisher $z$-transformation function $\text{tanh}^{-1}(\cdot)$ as $T_{F}(\cdot)$. Assuming each pair $(Y,X_{j})$ to be bivariate normally distributed, the corresponding transformed correlations are approximately normally distributed as well with known variance:
\
\begin{equation} \label{eq:fisher_transformation}
T_{F}(r_{j})\overset{ind}{\sim} N(T_{F}(\rho_{j}),\frac{1}{n-3}) \quad j=1,..,p
\end{equation}
Therefore, rescaling the quantities $T_{F}(r_{j})$ by $\sqrt{n-3}$ and referring to the results as variables $z_{j}$, we again find the formulation \eqref{eq:normal_means_problem} of the normal means problem, this time with unit variance:
\
\begin{equation} \label{eq:normal_means_problem2}
Z_{j}|\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \quad j=1,..,p
\end{equation}
Note that in thise cse the quantities of interest $\theta_{j}$ are equal to $\sqrt{n-3}\,T_{F}(\rho_{j})$. There are different ways to proceed with the selection in the normal means problem. Here, we consider several state-of-the-art approaches from both Bayesian and frequentist statistics (TODO: do we actually consider frequentist approaches here?)

We begin by explaining some Bayesian estimation methods to solving the normal means problem. They all have in common that a Bayesian regression model is fitted using some sparsifying prior on the regression coefficients, for example, as discussed in \cite{paper:dirichlet_laplace}, \cite{bhadra2017horseshoe+} and \cite{johnstone2004needles}. If these priors have continuous support, as is the case for the Dirichlet-Laplace (DL; cite) or horseshoe-like priors (cite), there is no canonical way to complete the selection since the posterior probability of an exactly zero signal is always zero. \cite{paper:dirichlet_laplace} suggest k-means clustering over the posterior samples using the DL prior, whereas \cite{bhadra2017horseshoe+} focus the analysis on the obtained shrinkage and the correct estimation of the sparse signals using the horseshoe+ prior (TODO: is this related to the former sentence?). In our experiments, we include the DL prior and the regularised horseshoe (RHS) prior \citep{paper:rhs}, which we think scales better to higher number of covariates as compared to other horseshoe-like priors. For the Bayesian regression models with these priors, we compare the posterior shrinkage and the correct estimation of the sparse signals with and without using a reference model as a noise filter. For this purpose, we consider the   average SSE (sum of square errors) and the average SESE (sum of expected square errors) as quantities measuring estimation accuracy. They are defined as
\
\begin{align}
\text{SSE}&=\frac{1}{n-3}\sum_{j=1}^{p}(\hat{\theta}_{j} - \theta^{0}_{j})^{2} \label{eq:SSE} \\
\text{SESE}&=\frac{1}{n-3}\sum_{j=1}^{p}\int_{\mathbb{R}}(\theta_{j}-\theta^{0}_{j})^{2}p(\theta_{j}|z_{1},..,z_{p})d\theta_{j} \\
&=\frac{1}{n-3}\sum_{j=1}^{p}\mathbb{E}_{\theta_{j}}[(\theta_{j}-\theta^{0}_{j})^{2}|z_{1},..,z_{p}] \label{eq:SESE}
\end{align}
where $\theta_{j}^{0}=\sqrt{n-3}\,T_{F}(\rho_{j})$ is the true value, $\hat{\theta}_{j}$ is the posterior median of each parameter $\theta_{j}$. We divide the errors by $(n-3)$ to have quantities independent of the number of statistical units. Note that the SESE takes the posterior uncertainty in account and thus also the shrinkage of the posterior distributions, whereas the SSE only measures the bias in the point estimates. Note that actually any kind of point estimate could be used to compute the SSE but here we follow \cite{paper:dirichlet_laplace} and use the posterior median. The results that we show later on in Section \ref{shrinkage-signal} are not sensible to such a choice.

We also take into account the posterior mean for the shrinkage factor ($\kappa_{j}$) for each parameter $\theta_{j}$, which for model \eqref{eq:normal_means_problem2} with a hierarchical shrinkage prior of the type (TODO: please motivate why looking at shrinkage factors is relevant)
\begin{align}
Z_{j}|&\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \qquad j=1,..,p \\
&\theta_{j}|\tau,\nu_{j}\overset{ind}{\sim}N(0,\nu_{j}^{2}\tau^{2})
\end{align}
is defined as
\begin{equation}
\kappa_{j}=\frac{1}{1+\tau^{2}\nu_{j}^{2}}.
\end{equation}
\
In our case, $\kappa_{j}$ can be computed as
\begin{equation}
\kappa_{j}^{\text{RHS}}=\frac{1}{1+\tau^{2}\lambda_{j}^{2}}, \qquad \kappa_{j}^{\text{DL}}=\frac{1}{1+\psi\phi_{j}^{2}\tau^{2}}
\end{equation}
respectively for the regularised horseshoe and the Dirichlet-Laplace priors. In the left hand-side expression, $\tau$ and $\lambda_{j}$ represent the global and local shrinkage parameters of the regularised horseshoe prior \cite[for further details see][]{paper:rhs}, whereas on the right hand-side, $\psi\phi_{j}^{2}\tau^{2}$ is the variance term of the Dirichlet-Laplace prior of parameter $\theta_{j}$ following the notation of \cite{paper:dirichlet_laplace}.

Within the set of complete selection procedures, we consider three 
Bayesian methods which are described in the following.
\cite{johnstone2004needles} use a spike-and-slab type prior given by a mixture of a delta distribution in zero and a heavy-tailed distribution \cite[see also][]{paper:spike_slab_mitchell}. When the latter is a Laplace distribution, they show an interesting thresholding property of the median estimator and, based on this observation, present a complete selection procedure using their prior. We refer to this method as EB.med. In our experiments, we not only include the EB.med on its own but also in combination with an additional control of the local false discovery rate (loc.fdr) as described in \cite{paper:efron, efron2012large}. Further, we device a simple selection procedure based on posterior credible intervals using the regularised horseshoe prior in which we select covariates whose 90\% credible intervals excludes zero (ci.90). All of these methods provide a complete selection procedure and we compare their performance with and without using a reference model on top of them. As criteria of selection performance, we evaluate the average false discovery rate (i.e., the ratio of the number of non-relevant selected features over the number of selected features) and the average sensitivity (i.e., the ratio of the number of relevant selected features over the total number of relevant features). We also provide a comparison of the stability of the selection by means of a stability measure proposed by \cite{paper:stability}. Table \ref{tab:comparison} summarises the methods used in out numerical experiments.

\begin{table}[tp]
\scriptsize
\centering
\begin{tabular}{l|l|l|l}
  \hline
 Method & Alias & Type & Quantities of interest \\ 
  \hline
 Regularised Horseshoe & RHS & Shrinkage prior & SSE, SESE, $\kappa$ \\
 Dirichlet-Laplace & DL & Shrinkage prior & SSE, SESE, $\kappa$ \\
 Local false discovery rate & loc.fdr & Complete selection & fdr, sensitivity, stability \\
 Median estimator & EB.med & Complete selection & fdr, sensitivity, stability \\
 Credible intervals & ci.90 & Complete selection & fdr, sensitivity, stability \\
 \hline
\end{tabular}
\caption{Summary of the shrinkage and selection methods used in the numerical experiments.}
\label{tab:comparison}
\end{table}
